Robinson et al. 2020 Supplemental Information: Comparing the Radiocarbon and Tree-ring Records for the Upland US Southwest

### 2020-01-18 analysis/paper/paper.Rmd
gpl3
binder

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

@@

# Summary of methods
Here, we directly compare two large collections of radiocarbon and tree-ring dates drawn from the Upland US Southwest --- roughly defined as the 'Four Corners' states of Arizona, Colorado, New Mexico, and Utah, but excluding the Phoenix and Tucson basins.

@@

## Study Area

@@

UUSW_counties <- 
  tigris::counties(cb = TRUE, 
                   class = "sf",
                   progress_bar = FALSE) %>% 
  sf::st_transform(four_corners_lcc_proj) %>% 
  # sf::st_contains(UUSW, sparse = FALSE)
  dplyr::filter(sf::st_intersects(suppressWarnings(sf::st_centroid(.)),
                                  UUSW %>% 
                                    sf::st_transform(four_corners_lcc_proj), 
                                  sparse = FALSE)) %>%
  dplyr::select(NAME,STATEFP) %>%
  dplyr::rename(County = NAME) %>%
  dplyr::left_join(tigris::states(class = "sf",
                                  progress_bar = FALSE) %>%
                     dplyr::select(STATEFP,STUSPS) %>%
                     dplyr::rename(State = STUSPS) %>%
                     sf::st_drop_geometry(),
                   by = "STATEFP") %>%
  dplyr::select(-STATEFP) %>%
  tidyr::unite(col = County, County, State, sep = ", ")

@@

## Datasets

## Preparing the radiocarbon data

@@

<!-- ```{r} -->
<!-- four_corners_lcc_proj <- "+proj=lcc +lat_1=37 +lon_0=-109.045225" -->

<!-- # Define the Bocinsky et al. 2016 study area -->
<!-- SWUS_bbox <- sf::st_bbox(c(xmin = -113, -->
<!--                            xmax = -105, -->
<!--                            ymin = 32, -->
<!--                            ymax = 38),  -->
<!--                          crs = 4326) %>% -->
<!--   sf::st_as_sfc() %>% -->
<!--   sf::st_set_crs(NA) %>% -->
<!--   sf::st_segmentize(dfMaxLength = 0.01) %>% -->
<!--   sf::st_set_crs(4326) %>% -->
<!--   sf::st_transform(four_corners_lcc_proj) -->

<!-- four_corners_states <- "./data-raw/statesp010g/statesp010g.gdb/" %>% -->
<!--   sf::st_read() %>% -->
<!--   dplyr::filter(NAME %in%  -->
<!--                   c("Arizona", -->
<!--                     "Colorado", -->
<!--                     "New Mexico", -->
<!--                     "Utah")) %>% -->
<!--   sf::st_transform(four_corners_lcc_proj) -->

<!-- swus_hillshade_500m <- readr::read_rds("./data-derived/swus_hillshade_500m.rds") %>% -->
<!--   raster::crop(four_corners_states) %>% -->
<!--   raster::mask(four_corners_states) -->

<!-- dendro_data <- sf::read_sf("./data-derived/dendro_data.geojson") %>% -->
<!--   sf::st_transform(four_corners_lcc_proj) %>% -->
<!--   dplyr::filter(Date_BP >= 550, -->
<!--                 Date_BP <= 1450) %>% -->
<!--   dplyr::group_by(Site_Number) %>% -->
<!--   dplyr::summarise(Date_BP = mean(Date_BP, na.rm = T), -->
<!--                    geometry = mean(geometry)) %>% -->
<!--   dplyr::mutate(`Date Type` = "Dendro") %>% -->
<!--   dplyr::select(-Site_Number) %>% -->
<!--   sf::st_centroid() -->

<!-- radiocarbon_data <- sf::read_sf("./data-derived/radiocarbon_data.geojson") %>% -->
<!--   sf::st_transform(four_corners_lcc_proj) %>% -->
<!--   dplyr::filter(Date_BP >= 550, -->
<!--                 Date_BP <= 1450) %>% -->
<!--   dplyr::group_by(Site_Number) %>% -->
<!--   dplyr::summarise(Date_BP = mean(Date_BP, na.rm = T)) %>% -->
<!--   sf::st_jitter(20000) %>% -->
<!--   dplyr::mutate(`Date Type` = "C14") %>% -->
<!--   dplyr::select(-Site_Number) %>% -->
<!--   sf::st_centroid() -->

<!-- dates <- rbind(dendro_data, -->
<!--                radiocarbon_data) -->

<!-- theme_map <- function(base_size = 6.5,  -->
<!--                       base_family = ""){ -->
<!--   ggplot2::theme_bw(base_size = base_size,  -->
<!--                     base_family = base_family) %+replace% -->
<!--     ggplot2::theme(axis.line = ggplot2::element_blank(), -->
<!--                    axis.text = ggplot2::element_blank(), -->
<!--                    axis.ticks = ggplot2::element_blank(), -->
<!--                    axis.title = ggplot2::element_blank(), -->
<!--                    panel.background = ggplot2::element_blank(), -->
<!--                    panel.border = ggplot2::element_blank(), -->
<!--                    panel.grid = ggplot2::element_blank(), -->
<!--                    panel.grid.major = ggplot2::element_line(colour = 'transparent'), -->
<!--                    panel.grid.minor = ggplot2::element_line(colour = 'transparent'), -->
<!--                    legend.key = element_blank(), -->
<!--                    legend.position = "top", -->
<!--                    # legend.justification = c(0,0), -->
<!--                    # legend.position = "top", -->
<!--                    # legend.position = c(0.5,1), -->
<!--                    # legend.background = ggplot2::element_blank(), -->
<!--                    # legend.key.width = unit(0.15,"in"), -->
<!--                    # legend.key.height = unit(0.15,"in"), -->
<!--                    # legend.text = element_text(size = ggplot2::rel(1)), -->
<!--                    plot.background = ggplot2::element_blank(), -->
<!--                    plot.margin = ggplot2::margin(t = 0, r = 0, b = 0, l = 0, unit = "npc") -->
<!--     ) -->
<!-- } -->

<!-- ggplot() +  -->
<!--   ggplot2::geom_sf(data = four_corners_states) + -->
<!--   ggplot2::geom_raster(data = Robinson2020::swus_hillshade_500m, -->
<!--                        mapping = ggplot2::aes(x = x, -->
<!--                                               y = y, -->
<!--                                               alpha = ID), -->
<!--                        na.rm = TRUE) + -->
<!--   ggplot2::scale_alpha(range = c(0.8, 0), -->
<!--                        na.value = 0, -->
<!--                        limits = c(0,255), -->
<!--                        guide = "none") + -->
<!--   ggplot2::geom_sf(data = four_corners_states, -->
<!--                    fill = NA) + -->
<!--   ggplot2::geom_sf(data = SWUS_bbox, -->
<!--                    fill = NA) + -->
<!--   ggplot2::geom_sf(data = dates, -->
<!--                    mapping = aes(shape = `Date Type`, -->
<!--                                  color = Date_BP), -->
<!--                    show.legend = "point") + -->
<!--   ggplot2::scale_shape( -->
<!--     name = 'Date Type', -->
<!--     solid = FALSE, -->
<!--     guide = ggplot2::guide_legend(position = "top", -->
<!--                                   direction = "horizontal", -->
<!--                                   title.position = "top", -->
<!--                                   label.position = "bottom") -->
<!--   ) + -->
<!--   viridis::scale_color_viridis( -->
<!--     name = 'Year BP', -->
<!--     limits = c(450,1450), -->
<!--     breaks = c(450,1450), -->
<!--     direction = -1, -->
<!--     guide = ggplot2::guide_colorbar(position = "top", -->
<!--                                     direction = "horizontal", -->
<!--                                     title.position = "top", -->
<!--                                     label.position = "bottom", -->
<!--                                     reverse = TRUE)) + -->
<!--   theme_map(16) -->

<!-- ggsave('./swus_sample_map.pdf', -->
<!--        height = 8, -->
<!--        width = 6) -->
<!-- ``` -->

>>>>

@@

# Summary of methods
Here, we directly compare two large collections of radiocarbon and tree-ring dates drawn from the Upland US Southwest.

@@

# Study Area

@@

UUSW_counties <- 
  tigris::counties(cb = TRUE, 
                   class = "sf",
                   progress_bar = FALSE) %>% 
  sf::st_transform(four_corners_lcc_proj) %>% 
  dplyr::filter(sf::st_intersects(suppressWarnings(sf::st_centroid(.)),
                                  UUSW %>% 
                                    sf::st_transform(four_corners_lcc_proj), 
                                  sparse = FALSE)) %>%
  dplyr::select(NAME,STATEFP) %>%
  dplyr::rename(County = NAME) %>%
  dplyr::left_join(tigris::states(class = "sf",
                                  progress_bar = FALSE) %>%
                     dplyr::select(STATEFP,STUSPS) %>%
                     dplyr::rename(State = STUSPS) %>%
                     sf::st_drop_geometry(),
                   by = "STATEFP") %>%
  dplyr::select(-STATEFP) %>%
  tidyr::unite(col = County, County, State, sep = ", ")

@@

The UUSW was geographically definied by @Bocinsky2016_SA as the region between 105--113ºW and 32--38ºN. Here, we refine that definition in two ways. First, since location data for our radiocarbon database are only resolved to the county level, we define the study region to include all counties whose geographic centroids fall within the region from @Bocinsky2016_SA. Second, we roughly exclude the Phoenix and Tucson basins by excluding Maricopa, Pinal, and Pima counties in Arizona. Figure \@ref(fig:study-area) shows the original UUSW from @Bocinsky2016_SA, and the revised study area used in this analysis.

```{r study-area}
theme_map <- function(base_size = 6.5,
                      base_family = ""){
  ggplot2::theme_bw(base_size = base_size,
                    base_family = base_family) %+replace%
    ggplot2::theme(axis.line = ggplot2::element_blank(),
                   axis.text = ggplot2::element_blank(),
                   axis.ticks = ggplot2::element_blank(),
                   axis.title = ggplot2::element_blank(),
                   panel.background = ggplot2::element_blank(),
                   panel.border = ggplot2::element_blank(),
                   panel.grid = ggplot2::element_blank(),
                   panel.grid.major = ggplot2::element_line(colour = 'transparent'),
                   panel.grid.minor = ggplot2::element_line(colour = 'transparent'),
                   legend.key = ggplot2::element_blank(),
                   legend.position = "top",
                   # legend.justification = c(0,0),
                   # legend.position = "top",
                   # legend.position = c(0.5,1),
                   # legend.background = ggplot2::element_blank(),
                   # legend.key.width = unit(0.15,"in"),
                   # legend.key.height = unit(0.15,"in"),
                   # legend.text = element_text(size = ggplot2::rel(1)),
                   plot.background = ggplot2::element_blank(),
                   plot.margin = ggplot2::margin(t = 0, r = 0, b = 0, l = 0, unit = "npc")
    )
}

four_corners_states <- 
  tigris::states(class = "sf",
                 progress_bar = FALSE) %>%
  dplyr::filter(NAME %in% c("Arizona", "Colorado", 
                            "New Mexico", "Utah")) %>%
  sf::st_transform(four_corners_lcc_proj)

ggplot2::ggplot() +
  ggplot2::geom_sf(data = four_corners_states) +
  ggplot2::geom_raster(data = robinson2020::swus_hillshade_500m,
                       mapping = ggplot2::aes(x = x,
                                              y = y,
                                              alpha = ID),
                       na.rm = TRUE) +
  ggplot2::scale_alpha(range = c(0.8, 0),
                       na.value = 0,
                       limits = c(0,255),
                       guide = "none") +
  ggplot2::geom_sf(data = four_corners_states,
                   fill = NA) +
  ggplot2::geom_sf(data = UUSW,
                   fill = NA) +
    ggplot2::geom_sf(data = UUSW_counties,
                   fill = NA) +
  # ggplot2::geom_sf(data = dates,
  #                  mapping = aes(shape = `Date Type`,
  #                                color = Date_BP),
  #                  show.legend = "point") +
  # ggplot2::scale_shape(
  #   name = 'Date Type',
  #   solid = FALSE,
  #   guide = ggplot2::guide_legend(position = "top",
  #                                 direction = "horizontal",
  #                                 title.position = "top",
  #                                 label.position = "bottom")
  # ) +
  # viridis::scale_color_viridis(
  #   name = 'Year BP',
  #   limits = c(450,1450),
  #   breaks = c(450,1450),
  #   direction = -1,
  #   guide = ggplot2::guide_colorbar(position = "top",
  #                                   direction = "horizontal",
  #                                   title.position = "top",
  #                                   label.position = "bottom",
  #                                   reverse = TRUE)) +
  theme_map(16)

# Datasets

# Preparing the radiocarbon data

@@

# Define the Bocinsky et al. 2016 study area
SWUS_bbox <- sf::st_bbox(c(xmin = -113,
                           xmax = -105,
                           ymin = 32,
                           ymax = 38),
                         crs = 4326) %>%
  sf::st_as_sfc() %>%
  sf::st_set_crs(NA) %>%
  sf::st_segmentize(dfMaxLength = 0.01) %>%
  sf::st_set_crs(4326) %>%
  sf::st_transform(four_corners_lcc_proj)

four_corners_states <- "./data-raw/statesp010g/statesp010g.gdb/" %>%
  sf::st_read() %>%
  dplyr::filter(NAME %in%
                  c("Arizona",
                    "Colorado",
                    "New Mexico",
                    "Utah")) %>%
  sf::st_transform(four_corners_lcc_proj)

swus_hillshade_500m <- readr::read_rds("./data-derived/swus_hillshade_500m.rds") %>%
  raster::crop(four_corners_states) %>%
  raster::mask(four_corners_states)

dendro_data <- sf::read_sf("./data-derived/dendro_data.geojson") %>%
  sf::st_transform(four_corners_lcc_proj) %>%
  dplyr::filter(Date_BP >= 550,
                Date_BP <= 1450) %>%
  dplyr::group_by(Site_Number) %>%
  dplyr::summarise(Date_BP = mean(Date_BP, na.rm = T),
                   geometry = mean(geometry)) %>%
  dplyr::mutate(`Date Type` = "Dendro") %>%
  dplyr::select(-Site_Number) %>%
  sf::st_centroid()

radiocarbon_data <- sf::read_sf("./data-derived/radiocarbon_data.geojson") %>%
  sf::st_transform(four_corners_lcc_proj) %>%
  dplyr::filter(Date_BP >= 550,
                Date_BP <= 1450) %>%
  dplyr::group_by(Site_Number) %>%
  dplyr::summarise(Date_BP = mean(Date_BP, na.rm = T)) %>%
  sf::st_jitter(20000) %>%
  dplyr::mutate(`Date Type` = "C14") %>%
  dplyr::select(-Site_Number) %>%
  sf::st_centroid()

dates <- rbind(dendro_data,
               radiocarbon_data)

theme_map <- function(base_size = 6.5,
                      base_family = ""){
  ggplot2::theme_bw(base_size = base_size,
                    base_family = base_family) %+replace%
    ggplot2::theme(axis.line = ggplot2::element_blank(),
                   axis.text = ggplot2::element_blank(),
                   axis.ticks = ggplot2::element_blank(),
                   axis.title = ggplot2::element_blank(),
                   panel.background = ggplot2::element_blank(),
                   panel.border = ggplot2::element_blank(),
                   panel.grid = ggplot2::element_blank(),
                   panel.grid.major = ggplot2::element_line(colour = 'transparent'),
                   panel.grid.minor = ggplot2::element_line(colour = 'transparent'),
                   legend.key = element_blank(),
                   legend.position = "top",
                   # legend.justification = c(0,0),
                   # legend.position = "top",
                   # legend.position = c(0.5,1),
                   # legend.background = ggplot2::element_blank(),
                   # legend.key.width = unit(0.15,"in"),
                   # legend.key.height = unit(0.15,"in"),
                   # legend.text = element_text(size = ggplot2::rel(1)),
                   plot.background = ggplot2::element_blank(),
                   plot.margin = ggplot2::margin(t = 0, r = 0, b = 0, l = 0, unit = "npc")
    )
}

ggplot() +
  ggplot2::geom_sf(data = four_corners_states) +
  ggplot2::geom_raster(data = Robinson2020::swus_hillshade_500m,
                       mapping = ggplot2::aes(x = x,
                                              y = y,
                                              alpha = ID),
                       na.rm = TRUE) +
  ggplot2::scale_alpha(range = c(0.8, 0),
                       na.value = 0,
                       limits = c(0,255),
                       guide = "none") +
  ggplot2::geom_sf(data = four_corners_states,
                   fill = NA) +
  ggplot2::geom_sf(data = SWUS_bbox,
                   fill = NA) +
  ggplot2::geom_sf(data = dates,
                   mapping = aes(shape = `Date Type`,
                                 color = Date_BP),
                   show.legend = "point") +
  ggplot2::scale_shape(
    name = 'Date Type',
    solid = FALSE,
    guide = ggplot2::guide_legend(position = "top",
                                  direction = "horizontal",
                                  title.position = "top",
                                  label.position = "bottom")
  ) +
  viridis::scale_color_viridis(
    name = 'Year BP',
    limits = c(450,1450),
    breaks = c(450,1450),
    direction = -1,
    guide = ggplot2::guide_colorbar(position = "top",
                                    direction = "horizontal",
                                    title.position = "top",
                                    label.position = "bottom",
                                    reverse = TRUE)) +
  theme_map(16)

ggsave('./swus_sample_map.pdf',
       height = 8,
       width = 6)
```
